Quantitative Benchmarking for Automatically Generated Conjectures

### 2018-04-26 TEBenchmarkPaper/article.tex
Fix a bunch of FIXMEs

% FIXME: Perhaps mention examples like "plus(x, 0) = 0" as motivating examples
% of things which are "simple" but which proof assistants still don't do
% automatically (reference Moa's use of TE for background lemma generation)

@@

% FIXME: We jump from conjectures to theorems. Maybe to prevent confusion we
% should talk more generally about "interesting properties" of a formal system,
% and use that as motivation for conjecturing and proving?

Given a logical theory, such as the theory of lists shown in
Figure~\ref{figure:list_theory}, we may want to find properties (such as
invariants) describing its behaviour. This could be for mathematical curiosity,
or due to the theory's importance in some domain. In particular, when dealing
with definitions in a software library, we may regard a set of such properties
as the code's \emph{specification}. Even those properties which only hold by
coincidence, rather than by design, may still be useful for \emph{optimising}
uses of this library, for example by rewriting expressions into equivalent forms
which require less time, memory, network usage, etc. For example, our theory of
lists turns out to satisfy the following invariant:

@@

Whilst these two expressions produce equivalent results, the form on the right
can execute both calls to $\Map$ in parallel (in a pure functional language, at
least), giving rise to the ``map/reduce'' paradigm for parallel programming.
Such use cases require the ability to \emph{discover} properties about some
particular definitions; this requires not only \emph{proving} their correctness,
but also \emph{posing} them as conjectures in the first place. This is a hard
problem in general, but presents opportunities for automation due to the
precise, symbolic nature of the domain.

@@

The task of automatically proving or disproving a \emph{given} conjecture has
been studied extensively in the field of Automated Theorem Proving (ATP). Less
attention has been paid to \emph{generating} such conjectures automatically,
either from a given set of definitions (the task we will focus on) or where the
definitions are also generated. The fields of Automated Theory Formation (ATF)
and Mathematical Theory Exploration (MTE) both tackle these generation tasks,
and we discuss their relationship in historical context in
$\S$\ref{related-work}.

A major challenge for both ATF and MTE is choosing how to narrow down
the set of generated conjectures to those deemed ``interesting'', since this is
an imprecise term with many different interpretations. For example, all existing
approaches agree that simple tautologies are ``uninteresting'', but differ when
it comes to more complex statements.  Colton \etal{} give a survey of tools for
theory formation and exploration, and their associated notions of
``interestingness'' for concepts and conjectures~\cite{colton2000notion}. Six
notions are identified, which are applied to conjectures as follows:

{\bf Empirical plausibility} checks whether a property holds across some
specific examples. This is especially useful for avoiding false conjectures,
without resorting to a full proof search.

{\bf Novelty} depends on whether a conjecture, or one isomorphic or more
general, has already been seen.

{\bf Surprisingness} of a conjecture is whether or not it is
``obvious'', for example if it is an instance of a tautology.

{\bf Applicability} depends on the number of models in which a
conjecture holds. The tool of \Bagai{} conjectures the non-existence of
objects, and hence favours statements with zero applicability. Other tools treat
applicability as a positive aspect: the more applicable the statement, the more
interesting it is.

{\bf Comprehensibility} depends on the complexity of a statement. Simpler
statements are considered more interesting, hence many of the tools explore
simpler/smaller statements before complex/larger ones, to more efficiently find
those which are interesting.

{\bf Utility} is the relevance or usefulness of a conjecture to the
user's particular task. For example, if we want to find optimising
rewrite rules such as equation~\ref{eq:mapreduce}, then utility would include
whether or not a conjecture justifies a rewrite rule, the difference in
resource usage of the expressions involved, and how common those
expressions are in real usage.

% FIXME: This is original work, so should be in a different section
\begin{table}
  \centering
  \begin{tabular}{ |l|l|c|c|c|c|c|c| }
    \hline
    \multirow{2}{*}{\textbf{Program}}                      &
    \multirow{2}{*}{\textbf{Conjecture Types}}             &
    \multicolumn{6}{c|}{\textbf{Interestingness Measures}} \\ \hhline{~~------}
    \tRow{          &                     & \iE & \iN & \iS & \iA & \iC & \iU}
    \tRow{AM        & \tIFF, \tIMP, \tNE  &   X &   X &   X &   X &   X &   X}
    \tRow{GT        & \tIFF, \tIMP, \tNE  &   X &   X &   X &   X &   X &   X}
    \tRow{Graffiti  & \tINE               &   X &   X &   X &     &   X &   X}
    \tRow{\Bagai{}  & \tNE                &     &   X &     &   X &   X &    }
    \tRow{HR        & \tIFF, \tIMP, \tNE  &   X &   X &   X &   X &   X &   X}
    \tRow{QuickSpec & \tEQ                &   X &   X &     &   X &   X &    }
    \tRow{Speculate & \tCON\ \tEQ / \tINE &   X &   X &     &   X &   X &    }
    \tRow{IsaCoSy   & \tEQ                &   X &   X &     &   X &   X &   X}
    \tRow{IsaScheme & \tEQ                &   X &   X &     &   X &   X &   X}
  \end{tabular}
  \caption{Classification of MTE tools from~\cite{colton2000notion}, extended
    to include four more recent tools. The interestingness measures are
    \iE{}mpirical plausibility, \iN{}ovelty, \iS{}urprisingness,
    \iA{}pplicability, \iC{}omprehensibility (low complexity) and \iU{}tility.}
  \label{table:colton}
\end{table}

% JUSTIFICATIONS
%
% IsaCoSy uses counter-example checking; this ensures empirical plausibility
% IsaCoSy uses constraints to avoid special-cases; this ensures novelty
% IsaCoSy uses utility, since there are some hard-coded patterns which are
% looked for
%
% Speculate uses a form of unification to ensure novelty, making sure one
% equation is not a special-case of another
% Speculate uses LeanCheck to enumerate values, looking for counterexamples
% Speculate uses testing, to ensure Empirical Plausibility
%
% QuickSpec uses QuickCheck to ensure empirical plausibility
% QuickSpec uses a congruence closure algorithm to ensure novelty
%
% IsaScheme uses utility, since it determines the "quality" of a definition
% based on how many theorems it appears in?
% IsaScheme uses novelty, using an equational rewrite system (Knuth-Bendix
% completion) to remove redundancies
% IsaScheme uses utility, since it focuses on the terms and definitions of a
% user's theory. However, don't all of them?
% IsaScheme uses counterexample checking for empirical plausibility.

% ONLY APPLIES TO PRECONDITIONS
% APPLICABILITY: ensure a conjecture holds in many models (or exactly zero, in
% the case of Bagai et al). Does the testing-based approach of QuickSpec and
% Speculate ensure that there's a model? After all, these are concrete values
% rather than e.g. implications of some abstract specification.

% FIXME: Give brief description and reference for each system in the text.
% We could include the more lengthy descriptions of IsaCoSy and QuickSpec here?

We summarise the criteria in Table 1 and state which were used in key
ATF/MTE tools. The entries for AM, GT, Graffiti, \Bagai{} and HR are based
on this survey (the latter is Colton's own tool). We extend this analysis to
some more recent tools in the remaining rows (based on our understanding of
their function): QuickSpec~\cite{QuickSpec},
Speculate~\cite{braquehais2017speculate},
IsaCoSy~\cite{Johansson.Dixon.Bundy:conjecture-generation}
and IsaScheme~\cite{MontanoRivas2011}.

The latter tools are described in more detail in $\S$\ref{section:mte},
but all use testing to check for counterexamples, which ensures results are
empirically plausible and applicable (conditions are satisfiable, types are
inhabited, etc.). They also use a small-to-large search order, ensuring more
easily-comprehensibile conjectures are explored first.

IsaCoSy ensures novelty using a constraint system to prevent special-cases being
generated, whilst the others use post-hoc filters based on unification and term
rewriting to achieve the same result. IsaCoSy is able to look for commonly
desirable patterns, such as commutativity and associativity, which increases the
utility of its output, whilst IsaScheme is completely based around such
pattern-instantiation.

This diversity of approaches makes it difficult to compare the existing
evaluations of these tools directly. In particular, evaluation methods created
for one tool might not make sense for another, due to assumptions made about
the algorithms' operation. For example, the novelty filters used by QuickSpec
and Speculate are applied to the whole output set, guaranteeing that no
special-cases will appear; yet we cannot assume this in the case of IsaCoSy,
since its constraint solver allows special cases if they're found \emph{before}
the general case (it does not go back and discard previous results).

With such ambiguous and varied goals, approaches and assessment criteria it is
difficult to compare MTE tools in a quantitative way, and hence to form some
measure of ``progress'' for the field. Our proposed benchmarking methodology
attempts to remedy this by setting a clear target, hopefully in line with the
myriad goals of existing researchers, whilst being generally applicable to
tools of this type.

\subsection{Existing Evaluations}
\label{sec:existing}

There are three aspects to consider when evaluating a tool for conjecture
generation:

% FIXME: How are these referred to/considered in the paper?
\begin{enumerate}
\item The quality of each individual conjecture, which we will refer to as their
  ``interestingness''.
\item The quality of the \emph{set} of generated conjectures, which is important
  for criteria like novelty, as well as to assess consistency.
\item Resources used by the tool, to find the balance struck between output
  quality and time taken.
\end{enumerate}

We make the assumption that each tool is intended to be \emph{general}, and
hence isn't hard-coded or fine-tuned to only perform well in a narrow domain.
Examples where this does \emph{not} hold include the natural number heuristics
of AM discussed in $\S$\ref{section:atf}, as well as machine learning models
trained on some specific dataset (such as the benchmark we describe).

Evaluation of new tools and techniques for conjecture generation tends to be
\emph{ad hoc} and hence difficult to compare, whilst comparative surveys (such
as Colton's) are descriptive and qualitative rather than providing direct
numerical measurements.

One quantitative approach which has gained traction is \emph{precision/recall
  analysis}. This has been applied to IsaCoSy, IsaScheme and HipSpec, the latter
providing a direct comparison of these three
tools~\cite{claessen2013automating}.\footnote{Unfortunately this comparison only
  reports prior results from the other tools, rather than reproducing them.
  This makes the numbers less comparable, since each tool was tested with
  slightly different definitions (e.g. whether or not the exponential and
  maximum functions were included).} To measure precision and recall we first
choose a set of definitions (such as those in figure \ref{figure:list_theory})
and a \emph{ground truth} to treat as the ``ideal'' set of conjectures we would
like a tool to produce for those definitions (e.g. this might include
equation~\ref{eq:mapreduce}). Each tool is run and their outputs are compared to
the ground truth. To score 100\% on precision and recall, all of the conjectures
which appear in the ground truth must be generated, and nothing else:

\begin{itemize}
\item \emph{Precision} is the proportion of a tool's output which appears in
  the ground truth. This penalises overly-liberal tools which output large
  numbers of conjectures in the hope that some turn out to be ``good''.
\item \emph{Recall} is the proportion of the ground truth which appears in the
  tool's output. This penalises overly-conservative tools which generate very
  few conjectures as a way to avoid ``bad'' ones.
\end{itemize}

Precision and recall act as quality tests for sets of conjectures, fulfilling
one of our requirements. They also provide a coarse quality assessment for
individual conjectures: either a conjecture appears in the ground truth set or
it doesn't.

The analyses in~\cite{claessen2013automating} took their definitions and ground
truths from the standard library of the Isabelle proof assistant: since the
library authors have gone to the effort of stating, proving and including these
theorems in every copy of their software, this is a good indication that they
are interesting properties of those definitions.

Tests were performed with a theory of natural numbers and a theory of lists (the
definitions shown in figure~\ref{figure:list_theory}), %FIXME: add page ref
but it is difficult to
draw general conclusions from only two datapoints. These theories were also
quite small, the natural number theory
from~\cite{Johansson.Dixon.Bundy:conjecture-generation} only containing 4
definitions and 12 theorems.

Despite these limitations we believe that this is a viable approach to drawing
meaningful, objective conclusions about such tools; but only with larger data
sets can we judge performance in different, especially \emph{novel}, domains
(which, after all, is the purpose of such tools). In addition, we might be
concerned that such ``popular'' theories may be unrepresentative of typical
usage. For example, we would not expect the library functions in a typical
verification problem to be as mathematically rich as number theory.

@@

Our main contribution is a benchmarking methodology, shown in figure
\ref{figure:flow_chart}, which both generates a large definition/ground-truth
corpus, and provides a scalable, statistical approach to evaluating MTE tools
using this corpus. We follow a precision/recall approach similar to prior work,
with the main difference being the source of definitions and ground truths: we
take existing problem sets designed for automated theorem proving, and adapt
their content for use in the conjecture generation setting.

@@

\begin{itemize}
\item For each problem, there should be a clear distinction between the theorem
  to be proved and the definitions involved, such that the two can be easily and
  meaningfully separated. This rules out problem sets like those of
  SMT-COMP~\cite{barrett2005smt}, where many problems involve uninterpreted
  functions, whose behaviour is \emph{implicit} in the logical structure of the
  theorem statement but not separately \emph{defined}.
\item Definitions should be translatable into a form suitable for the MTE tools
  under study. Our application in $\S$\ref{sec:application} requires Haskell and
  Isabelle translations, and also benefits from having definitions be strongly
  typed.
\item The problem set should be relevant to the desired domain. Our focus on
  functional programming requires higher-order functions and inductively defined
  datatypes, which rules out first-order languages/logics (such as
  TPTP~\cite{sutcliffe2009tptp}).
\item Since it will act as our ground truth, the problem set should
  \emph{ideally} contain every ``interesting'' conjecture involving its included
  definitions. More \emph{realistically}, we should aim for each definition to
  appear in \emph{many} theorem statements; rather than each problem having
  unique definitions. % FIXME: Reword: make clear that non-membership means
  % uninteresting, and hence the things we *don't* include are just as important
  % as the things we do
\item The problem set should be as large as possible, for robustness of the
  resulting statistics. % FIXME: Prioritising quantity over quality?
\end{itemize}

@@

\subsection{Sampling}

@@

\subsection{Evaluation}

@@

    % Preparation
    \draw
      let \p{thm} = (thm.east),
          \p{in}  = (in.north),
          \p{rec} = (rec.south east)
       in [decorate,decoration={brace, amplitude=10pt}, xshift=0.5cm]
       (\x{rec}, \y{in}) -- (\x{rec}, \y{thm})
         node [black, midway, right, xshift=0.3cm] {Preparation};

    % Sampling
    \draw
      let \p{thm} = (thm.east),
          \p{gt}  = (sthm.east),
          \p{rec} = (rec.south east)
       in [decorate,decoration={brace, amplitude=10pt}, xshift=0.5cm]
       (\x{rec}, \y{thm}) -- (\x{rec}, \y{gt})
         node [black, midway, right, xshift=0.3cm] {Sampling};

    % Evaluation
    \draw
      let \p{sthm} = (sthm.east),
          \p{rec}  = (rec.south east)
       in [decorate,decoration={brace, amplitude=10pt}, xshift=0.5cm]
       (\x{rec}, \y{sthm}) -- (\x{rec}, \y{rec})
         node [black, midway, right, xshift=0.3cm] {Evaluation};
  \end{tikzpicture}
  \caption[]{High-level view of our benchmarking methodology, showing
    \begin{tikzpicture}
      \node [rectangle, text centered, draw=black]{processes};
    \end{tikzpicture},
    \begin{tikzpicture}
      \node [rectangle, rounded corners, text centered, draw=black]{data};
    \end{tikzpicture} and
    \begin{tikzpicture}
      \node [trapezium, trapezium left angle=70, trapezium right angle=110,
             text centered, draw=black]{inputs};
  \end{tikzpicture}}
  \label{figure:flow_chart}
\end{figure}

@@

IsaCoSy is a tool for the Isabelle proof assistant, mostly written in Standard
ML. We took IsaCoSy from version 2015.0.3 of the IsaPlanner project, and ran it
with the 2015 version of Isabelle. The following issues had to be overcome to
make our benchmark applicable to IsaCoSy:

@@

%IsaCoSy
The IsaCoSy system (Isabelle Conjecture Synthesis) conjectures equations by
enumerating expressions involving a given set of (typed) constants and free
variables (a \emph{signature}). A constraint solver forbids certain
(sub)expressions from being synthesised, and these constraints are extended
whenever a new conjecture is generated, to avoid generating any special-cases of
this conjecture in the future. Conjectures are tested by looking for
counterexamples and, if none are found, sent to IsaPlanner which attempts to
prove them.

%QuickSpec
QuickSpec emerged from work on the QuickCheck software testing framework for the
Haskell programming language~\cite{claessen2011quickcheck}, and has been used to
generate conjectures for the subsequent
HipSpec~\cite{Claessen_hipspec:automating} and Hipster~\cite{Hipster} MTE
systems. Like IsaCoSy, a signature is provided and terms are enumerated. Rather
than a constraint system, QuickSpec version 1 collects expressions of the same
type into equivalence classes, assuming them to be equal. QuickCheck is used to
test this assumption by instantiating each variable with a randomly chosen
value, evaluating the resulting expressions and comparing them. Any equivalence
class whose elements don't compare equal are split up to separate those
observably-different expressions, and the process is repeated with new random
values.

After 500 rounds of testing, any expressions still sharing the same equivalence
class are conjectured to be equal for all values of their variables. Finally, a
congruence closure algorithm is applied to these equations, to remove
redundancies and special-cases.

>>>>

@@

Given a logical theory, such as the theory of lists shown in
Figure~\ref{figure:list_theory}, we may want to find properties (e.g.
invariants) describing its behaviour. This could be for mathematical curiosity,
or due to the theory's importance in some domain. In particular, when dealing
with definitions in a software library, we may regard a set of such properties
as the code's \emph{specification}. Even those properties which only hold by
coincidence, rather than by design, may still be useful for \emph{optimising}
uses of this library, by rewriting expressions into equivalent forms which
require less time, memory, network usage, etc. For example, our theory of lists
turns out to satisfy the following invariant:

@@

Whilst these two expressions produce equivalent results, the form on the right
can execute both calls to $\Map$ in parallel (in a pure functional language, at
least); a fact which is exploited by the ``map/reduce'' parallel programming
paradigm.

Such use cases require the ability to \emph{discover} true properties of some
particular definitions; this requires not only \emph{proving} their correctness,
but also \emph{posing} them as conjectures in the first place. This is a hard
problem in general, but presents opportunities for automation due to the
precise, symbolic nature of the domain. These processes can also feed into each
other, since proven properties can be used as lemmas for subsequent proofs. As
an example, proving Equation~\ref{eq:mapreduce} may require ``background
knowledge'' which is tedious to produce by hand, such as
$\forall x. \Append(x, \Nil) = x$.

@@

The task of automatically proving or disproving a \emph{given} conjecture has
been studied extensively in the field of Automated Theorem Proving (ATP). Less
attention has been paid to \emph{generating} such conjectures automatically,
either from a given set of definitions (the task we will focus on) or where the
definitions are also generated. This research has been pursued under the names
of Automated Theory Formation (ATF) and Mathematical Theory Exploration (MTE);
we will use the MTE terminology, and discuss their relationship further in
$\S$\ref{sec:related-work}.

A major challenge when generating conjectures is choosing how to narrow down the
resulting set to only those deemed ``interesting'', since this is an imprecise
term with many different interpretations. For example, all existing approaches
agree that simple tautologies are ``uninteresting'', but differ when it comes to
more complex statements.  Colton \etal{} give a survey of tools for theory
formation and exploration, and their associated notions of
``interestingness'' for concepts and conjectures~\cite{colton2000notion}. Six
notions are identified, which are applied to conjectured properties as follows:

{\bf Empirical plausibility} checks whether a property holds across some
specific examples. This is especially useful for avoiding falsehoods, without
resorting to a deductive proof search.

{\bf Novelty} depends on whether the property, or one isomorphic or more
general, has already been seen.

{\bf Surprisingness} of a property is whether or not it is ``obvious'', for
example if it is an instance of a tautology.

{\bf Applicability} depends on the number of models in which a
property holds. High applicability makes a property more interesting, but so
does zero applicability (i.e. non-existence).

{\bf Comprehensibility} depends on the syntactic complexity of the property's
statement. Simpler statements are considered more interesting (which favours
tools adopting a small-to-large search order).

{\bf Utility} is the relevance or usefulness of a property to the user's
particular task. For example, if we want to find optimisations like
Equation~\ref{eq:mapreduce}, then utility would include whether the property
allows rewriting, the difference in resource usage of the expressions involved,
and how common those expressions are in real usage.

Attempting to \emph{measure} such qualities is difficult, and having too many
measures complicates comparisons. System developers have employed a more
practical alternative in their evaluations, which is to perform
\emph{precision/recall} analysis against a \emph{ground-truth}. This requires
choosing a set of definitions (such as those in Figure~\ref{figure:list_theory})
and a set of properties (the ground truth) which represent the ``ideal'' result
of conjecture generation for those definitions (e.g. this might include
Equation~\ref{eq:mapreduce}). To analyse the quality of a tool's conjectures, we
run it on these chosen definitions and compare its output to the ground truth:

\begin{itemize}
\item \emph{Precision} is the proportion of a tool's output which appears in
  the ground truth. This penalises overly-liberal tools which output a large
  number of properties in the hope that some turn out to be ``good''.
\item \emph{Recall} is the proportion of the ground truth which appears in the
  tool's output. This penalises overly-conservative tools which generate very
  few properties as a way to avoid ``bad'' ones.
\end{itemize}

To score 100\% on precision and recall, all of the properties which appear in
the ground truth must have been conjectured, and nothing else. This gives us a
simple method to evaluate and compare tools without requiring a general solution
to the question of what is ``interesting''; although we must still decide what
to put in the ground truth set, and what to leave out, for each measured set of
definitions.

The precision and recall of three MTE tools are compared
in~\cite{claessen2013automating}:
IsaCoSy~\cite{Johansson.Dixon.Bundy:conjecture-generation},
IsaScheme~\cite{MontanoRivas2011} and
HipSpec~\cite{Claessen_hipspec:automating}.\footnote{Unfortunately this
  comparison only reports prior results for each tool, rather than reproducing
  them. This makes the numbers less comparable, since each tool was tested with
  slightly different definitions (e.g. whether or not the exponential and
  maximum functions were included).} The definitions and ground truths for all
three were taken from the standard library of the Isabelle proof assistant:
since the library authors have gone to the effort of stating and proving these
properties, and including them with every copy of their software, this is a good
indication that they are interesting. Two sets of definitions were used: a
theory of natural numbers and of lists (which we show in
Figure~\ref{figure:list_theory} on page~\pageref{figure:list_theory}). With only
two datapoints containing only a few definitions and properties (e.g. the
natural number theory contained 4 functions with 12 properties), it is difficult
to draw general conclusions; yet despite these limitations we believe that this
is a viable method to performing meaningful, objective analyses and comparisons
for such tools. It is only with larger data sets that we can judge performance
in different, especially \emph{novel}, domains (which, after all, is the purpose
of MTE tools). In addition, we might be concerned that such ``popular'' theories
as the natural numbers may be unrepresentative of typical usage. For example, we
would not expect the library functions in a typical verification problem to be
as mathematically rich as number theory.

\subsection{Existing Tools}
\label{sec:existing-tools}

Since our benchmarking methodology builds on the precision/recall analysis
described in $\S$\ref{sec:te}, we decided to use two of the same tools, IsaCoSy
and QuickSpec (the conjecture generating component of HipSpec), as our
demonstration. Other notable tools, including IsaScheme, are described under
related work in $\S$\ref{sec:related-work}.

IsaCoSy (Isabelle Conjecture Synthesis) is written for the Isabelle proof
assistant, mostly in Standard
ML~\cite{Johansson.Dixon.Bundy:conjecture-generation}. It conjectures equations
by enumerating expressions involving a given set of (typed) constants and free
variables (a \emph{signature}). A constraint solver forbids certain
(sub)expressions from being synthesised, and these constraints are extended
whenever a new property is conjectured, to avoid generating any special-cases of
this property in the future. Conjectures are tested by looking for
counterexamples and, if none are found, sent to IsaPlanner which attempts to
prove them.

QuickSpec emerged from work on the QuickCheck software testing framework for the
Haskell programming language~\cite{claessen2011quickcheck}. As well as
generating conjectures for HipSpec (also a Haskell tool) it has also been
applied to Isabelle via the Hipster tool~\cite{Hipster}. Like IsaCoSy, QuickSpec
takes a signature and enumerates well-typed terms, collecting together those of
the same type into equivalence classes, assuming them to be equal. QuickCheck is
then used to find counterexamples to this assumption, which it does by randomly
instantiating the variables and comparing the resulting closed expressions. Any
equivalence class whose elements don't compare equal are split up, and the
process is repeated with new random values.

After 500 rounds of testing, any expressions still sharing the same equivalence
class are conjectured to be equal for all values of their variables. Rather than
using a constraint system to prevent redundancies (such as special-cases of
other properties), QuickSpec instead passes its output through a congruence
closure algorithm to achieve the same effect.

@@

Our main contribution is a benchmarking methodology, shown in
Figure~\ref{figure:flow_chart} on page~\pageref{figure:flow_chart}, which both
generates a large definition/ground-truth corpus, and provides a scalable,
statistical approach to evaluating MTE tools using this corpus. We follow a
precision/recall approach similar to prior work, with the main difference being
the source of definitions and ground truths: we take existing problem sets
designed for automated theorem proving, and adapt their content for use in the
conjecture generation setting.

@@

\begin{itemize}
\item For each problem, there should be a clear distinction between the theorem
  to be proved and the definitions involved, such that the two can be easily and
  meaningfully separated. This rules out problem sets like those of
  SMT-COMP~\cite{barrett2005smt}, where many problems involve uninterpreted
  functions, whose behaviour is \emph{implicit} in the logical structure of the
  theorem statement but not separately \emph{defined}.
\item Definitions should be translatable into a form suitable for the MTE tools
  under study. Our application in $\S$\ref{sec:application} requires Haskell and
  Isabelle translations, and also benefits from having definitions be strongly
  typed.
\item The problem set should be relevant to the desired domain. Our focus on
  functional programming requires higher-order functions and inductively defined
  datatypes, which rules out first-order languages/logics (such as
  TPTP~\cite{sutcliffe2009tptp}).
\item The problem set should ideally contain \emph{every} ``interesting''
  property involving its included definitions, since non-membership in the
  ground truth will be treated as being ``uninteresting''. More realistically,
  we should at least aim for each definition to have \emph{multiple} properties;
  rather than a mixture of mostly unrelated properties.
\item Larger problem sets are preferable, all else being equal, as they give
  more robust statistics.
\end{itemize}

@@

\subsection{Sampling}
\label{section:sampling}

@@

\subsection{Evaluation}
\label{section:evaluation}

@@

    % Preparation
    \draw
      let \p{thm} = (thm.east),
          \p{in}  = (in.north),
          \p{rec} = (rec.south east)
       in [decorate,decoration={brace, amplitude=10pt}, xshift=0.5cm]
       (\x{rec}, \y{in}) -- (\x{rec}, \y{thm})
         node [black, midway, right, xshift=0.3cm]
              {Preparation $\S$\ref{section:prep}};

    % Sampling
    \draw
      let \p{thm} = (thm.east),
          \p{gt}  = (sthm.east),
          \p{rec} = (rec.south east)
       in [decorate,decoration={brace, amplitude=10pt}, xshift=0.5cm]
       (\x{rec}, \y{thm}) -- (\x{rec}, \y{gt})
         node [black, midway, right, xshift=0.3cm]
              {Sampling $\S$\ref{section:sampling}};

    % Evaluation
    \draw
      let \p{sthm} = (sthm.east),
          \p{rec}  = (rec.south east)
       in [decorate,decoration={brace, amplitude=10pt}, xshift=0.5cm]
       (\x{rec}, \y{sthm}) -- (\x{rec}, \y{rec})
         node [black, midway, right, xshift=0.3cm]
              {Evaluation $\S$\ref{section:evaluation}};

  \end{tikzpicture}
  \caption[]{High-level view of our benchmarking methodology, showing
    \begin{tikzpicture}
      \node [rectangle, text centered, draw=black]{processes};
    \end{tikzpicture},
    \begin{tikzpicture}
      \node [rectangle, rounded corners, text centered, draw=black]{data};
    \end{tikzpicture} and
    \begin{tikzpicture}
      \node [trapezium, trapezium left angle=70, trapezium right angle=110,
             text centered, draw=black]{inputs};
  \end{tikzpicture}}
  \label{figure:flow_chart}
\end{figure}

@@

We took IsaCoSy from version 2015.0.3 of the IsaPlanner project, and ran it
with the 2015 version of Isabelle. The following issues had to be overcome to
make our benchmark applicable to IsaCoSy:

@@
